---
title: "prediction_exploration - Akash"
output: html_document
author: "Akash Mahajan (akashmjn@stanford.edu), Raunaq Rewari (raunaq@stanford.edu)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
sapply(c('ggplot2','data.table','tidyverse','DT','gridExtra','forecast','xts'),require,character.only=TRUE)
```

## Loading in data after all filtering and transformations 

Loading of data. Generating a dataset containing just the top 10 tracks. 

```{r dataoverview1, echo=FALSE}

# loading in script files
source("src_akash.R")

# read in only the top 20 regions 
dt = fread("data_t20regions.csv")  # faster than read.csv 
dt[,Date:=as.Date(Date)]
data = data.frame(dt)
data$Date = as.Date(data$Date)


# create train, test classification datasets 
posThreshold = 20
dtDaysInTop20 = dt[Position<=posThreshold,.(DaysInTop20=.N,
                                            MeanStreams=mean(MeanStreams))
                   ,by=.(URL,Region,TSID)]
dtDaysInTop20[,Label:=TRUE]
# read in song_features data
song_feats = fread("song_feats.csv")
dtClassification = merge(song_feats,dtDaysInTop20[,.(URL,MeanStreams,Label)],
                         by="URL",all.x = TRUE)
dtClassification = dtClassification[,by=URL]

ggplot(dtDaysInTop20)+geom_histogram(aes(x=DaysInTop20))+facet_wrap(~Region)

# # creating classification data labels
# posThreshold = 20
# dt[Position<=posThreshold,.N,by=.(URL,Region)]

# correlations of top tracks - auto, cross with regions


```

## Trend Analysis

```{r trends1, echo=FALSE}

top5Obj = getTopNTracks(dt,5)
dtTop5 = top5Obj[[1]]

```


## Baseline 

A baseline for this task would be to just predict the mean of the data.  

```{r exploration, echo=FALSE}

# Filtering down data to only time series of a minium duration
# Giving a unique ID to a (Track,Region) pair. (this corresponds to one time series)
tsFilterStat = filterValidTS(dt,28)
dtFilteredTS = tsFilterStat[[1]]
dtTSDurations = tsFilterStat[[2]]

## Making a dataset of only top 5 songs 
nTop = 5
topStat = getTopNTracks(dtFilteredTS,5)
dtTop = topStat[[1]]

# filter and only keep tracks that have made a rise in the charts
# i.e. min Position of track is less that threshold at some point 

regionVec = c('gb','se','us')

ggplot(dtTop[Region%in%regionVec])+
  geom_line(aes(x=Date,y=Streams,group=TrackName,alpha=0.2,color=Position))+
  facet_wrap(~Region,scale="free_y",nrow=length(regionVec))

```

## Approach

For an initial simple analysis, we choose just the GB region with about 41 tracks that have made it to the top 5. We randomly set aside (25\%) ~ 10 tracks that we will not use during our model building purposes. 

We plan to explore the use of time series models for forecasting these streams. We make use of an ARMA model

\[
X_t - \alpha_1X_{t-1}, \cdots, -\alpha_pX_{t-p} = \epsilon_t +\theta_1\epsilon_{t-1},\cdots,+\theta_q\epsilon_{t-q} 
\]

Model selection (between differently parameterized models), is done using in-sample estimates of the generalization error, namely the AIC. 

```{r exploration, echo=FALSE}

dtTopGB = dtTop[Region=='gb']
gbTopTSIDs = unique(dtTopGB$TSID)

trainTSIDs = gbTopTSIDs[c(1,3)]

dtTS = dtTopGB[TSID==trainTSIDs[2],.(Streams,Date)]
trainTS1 = xts(dtTS[,Streams],order.by = dtTS[,Date])

tsModel = auto.arima(trainTS1)

summary(tsModel)

ggplot(dtTS)+
  geom_line(aes(x=Date,y=Streams))+
  geom_line(aes(x=Date,y=tsModel$fitted),color='red')

```



```{r exploration, echo=FALSE}

```